---
title: "Open and clean WMO data"
author: "Jonathan Bourne"
date: "11. oktober 2015"
output: html_document
---

This document will contain the code that can be used to open and the WMO files and extract the data

summaries to take mean, varience total number

Load packages
```{r}

packages <- c("dplyr", "tidyr", "magrittr", "openxlsx", "caret")

lapply(packages, library, character.only = TRUE)


file.loc <-file.path("~","Benchmarking", "PA consulting", "2015", "Statkraft") #file location may be computer specific depending on network
setwd(file.loc)
```

#Extract relevant sheets as data frames

extraction functions
```{r}

powerplant <- function(filename, company) {
  power_plant_level <- read.xlsx(filename, sheet = "T Part A - Power plant level", startRow = 12, colNames = FALSE )[,1:26]
  
  names(power_plant_level) <- read.xlsx(filename, sheet = "T Part A - Power plant level" ,
                                        startRow = 8, rows = 8:9)[,-c(1,27,28)]%>% 
    names  %>% gsub("/|\\-|\\?", "",.)
  power_plant_level %<>% filter(!is.na(Station.ID))
   power_plant_level$Station.ID <- paste(company, power_plant_level$Station.ID, sep = "" )
 names(power_plant_level) <- gsub("\\?","", names(power_plant_level))
  power_plant_level 
}

gen_spec <- function(filename, company){
gen_spec <- read.xlsx(filename, sheet = "T part B - Spec. generator" , startRow = 12, colNames = FALSE )[,1:16]

names(gen_spec) <-  read.xlsx(filename, sheet = "T part B - Spec. generator" , 
                             startRow = 8, rows = 8:9)[,2:17]%>% 
                    names %>% gsub("\\?|\\/|\\(|\\)", "",.) #the ? are being read as Regex and so are removed from the headings, they could be escaed but are superflous anyway.
gen_spec %<>% filter(!is.na(Station.ID))
gen_spec$Station.ID <- paste(company, gen_spec$Station.ID, sep = "" )
catag <-c("Turbine.type","Reversible.pump.turbine","Turbine.with.2.or.more.runners","Water.cooled.windings")
gen_spec %<>%mutate_each_(funs(factor),catag) %>% 
          group_by(Station.ID) %>% 
          mutate(specID = 1, specID = paste(Station.ID,"_",cumsum(specID),sep="")) %>% 
  ungroup  
gen_spec
}

dam_spec <- function(filename, company) {
    dam_spec <- read.xlsx(filename, sheet = "T part C - Spec. reservoir, EBI", startRow = 12, colNames = FALSE )[,2:11]

    names(dam_spec) <- read.xlsx(filename, sheet = "T part C - Spec. reservoir, EBI", startRow = 8, rows = 8:9)[,2:11]%>% 
      names %>% gsub("\\?|\\/|\\(|\\)", "",.) #the ? are being read as Regex and so are removed from the headings, they could be escaed but are superflous anyway.
    dam_spec %<>% filter(!is.na(Station.ID))
    dam_spec$Station.ID <- paste(company, dam_spec$Station.ID, sep = "" )
    names(dam_spec)[6] <- "Distance.from.roads"
    catag <- c("Object.ID","Dam.or.Riverstream.intake.EBI","Altitude.above.sea.level"  ,"Climate.zone" ,"Dam.classification")
    dam_spec %<>% mutate_each_(funs(factor),catag) %>% 
          group_by(Station.ID) %>% 
          mutate(specID = 1, specID = paste(Station.ID,"_",cumsum(specID),sep="")) %>% 
          ungroup 
dam_spec
}

gates_spec <- function(filename, company) {
      gates_spec <- read.xlsx(filename, sheet = "T part D - Spec. gates above gr", startRow = 12, colNames = FALSE )[,1:9]

      names(gates_spec) <- read.xlsx(filename, sheet = "T part D - Spec. gates above gr", 
                               startRow = 8, rows = 8:9)[,2:10]%>% names %>% gsub("\\?|\\/", "",.)
      gates_spec %<>% filter(!is.na(Station.ID))
      gates_spec$Station.ID <- paste(company, gates_spec$Station.ID, sep = "" )
#names(gates_spec) <- gsub("\\?","", names(gates_spec))
      gates_spec$Object.ID <- gates_spec$Object.ID[sub(",", "", gates_spec$Object.ID)] # commas were causing problems with the csv import, even though they were strings.
      catag <- c("SectorRadial.gate","Flood.gate")
      gates_spec %<>% mutate_each_(funs(factor),catag) %>% 
          group_by(Station.ID) %>% 
          mutate(specID = 1, specID = paste(Station.ID,"_",cumsum(specID),sep="")) %>% 
          ungroup  
gates_spec
}

valves_spec <- function(filename, company) {
  valves_spec <- read.xlsx(filename, sheet = "T part E - Spec. valves", startRow = 12, colNames = FALSE )[,1:5]

names(valves_spec) <- read.xlsx(filename, sheet = "T part E - Spec. valves", startRow = 8, rows = 8:9)[,2:6]%>% names %>% gsub("\\?", "",.) #the ? are being read as Regex and so are removed from the headings.
valves_spec %<>% filter(!is.na(Station.ID)) #prevents loads of empty lines being imported
valves_spec$Station.ID <- paste(company, valves_spec$Station.ID, sep = "" )
# valves does not have any catagorical values to rollup
valves_spec %<>%  group_by(Station.ID) %>% 
                  mutate(specID = 1, specID = paste(Station.ID,"_",cumsum(specID),sep="")) %>%
                  ungroup 
valves_spec
}

rollup <- function(dataframe){
        dataframe %>% 
              group_by(Station.ID) %>% 
              mutate(count = n()) %>% 
              summarise_each_(funs(mean, max, min, sum, var), 
                  names(.)[sapply(., class)=="numeric"|sapply(., class)=="integer"])
}

catagoricals <- function(data) {
data <-  data %>% select(1,length(.),which(sapply(., class)=="factor"))

#split out catagorical values and replace rogue J's with Y remove NA values
data2 <- data %>% gather(variable, class, -c(Station.ID,specID))
data2$class <- sub("^J$", "Y", data2$class)
data2 %<>%  mutate(present = 1, variable = paste(variable, "_",class, sep="")) %>%
            select(-class) %>% 
            spread(variable, present, fill = 0) %>% select(-specID) %>%
  mutate(x_NA = 1) #this additional column prevents a bug on the next command which is if there area no columns ending in _NA all columns are removed.

data2 %<>% select(-ends_with("_NA"))

##Aggregate up all the values into a single powerplant
data2 %<>% group_by(Station.ID) %>% summarise_each(funs(sum))
}

aggregator <- function(data) {
  if (length(which(sapply(data, class)=="factor"))!=0){  
  rolled <- rollup(data)
  catag <- catagoricals(data)
  data <- full_join(rolled, catag, by = "Station.ID")
  } else {data <- rollup(data)}
    data
}

combinatron <- function(filename, comapanyname){
plantspec <- powerplant(filename, comapanyname) %>% aggregator
genspec <- gen_spec(filename, comapanyname) %>% aggregator
damspec <- dam_spec(filename, comapanyname) %>% aggregator
gatespec <- gates_spec(filename, comapanyname) %>% aggregator
valvespec <- valves_spec(filename, comapanyname) %>% aggregator
plantspec %<>% full_join(., genspec ,by="Station.ID") %>% 
                full_join(., damspec ,by="Station.ID") %>%
                full_join(., gatespec ,by="Station.ID") %>%
                full_join(.,valvespec,by="Station.ID")
plantspec
}

```

Load and combine files
```{r}
filename <- "Data Gathering Sheet 2015 V2.xlsx"
company <- "Statkraft"
triplex <- combinatron(filename, company)


stat_names <- c("statkraft_powerplant", "statkraft_genspec", "statkraft_damspec", "statkraft_gatespec", "statkraft_valvespec")

eco <- combinatron("ECO 2014.xlsx", "ECO")

eidsiva <- combinatron("Eidsiva 2013 DD1 (31.04.2014).xlsx","Eidsiva")
filename2 <- "Eidsiva 2013 DD1 (31.04.2014).xlsx"
company2 <- "Eidsiva"

plant_spec <- powerplant(filename, company) %>%aggregator
genspec <- gen_spec(filename, company) %>%aggregator
damspec <- dam_spec(filename, company)%>%aggregator
gatespec <- gates_spec(filename, company) %>%aggregator
valvespec <- valves_spec(filename, company)%>%aggregator

#filename <- filename2
#company <- company2
plant_spec2 <- powerplant(filename2, company2) %>%aggregator
genspec2 <- gen_spec(filename2, company2)%>%aggregator
damspec2 <- dam_spec(filename2, company2) %>%aggregator
gatespec2 <- gates_spec(filename2, company2)  %>%aggregator
valvespec2 <- valves_spec(filename2, company2)  %>%aggregator

test <- bind_rows(triplex, eidsiva, eco) #
test[is.na(test)] <- 0
test %<>%  mutate_each(funs(numeric),names(.)[sapply(., class)=="integer"])

cortest <-cor(test[,-1])
hc = findCorrelation(cortest, cutoff=0.8)
names(.)[sapply(., class)=="numeric"|sapply(., class)=="integer"]
```


Remove empty variables or nearly empty variables
```{r}

```


Remove Highly correlated variables
```{r}

```

Add Cost and WMO data
```{r}

```

Model WMO accuracy
```{r}

```

Create variety of statistical models
```{r}

```


```{r}
names(genspec2)[!(names(genspec2) %in% names(genspec))]
names(plant_spec2)[!(names(plant_spec2) %in% names(plant_spec))]
names(triplex)[!(names(eidsiva) %in% names(triplex))]

x1 <- bind_rows(genspec, genspec2)
x2 <- bind_rows(plant_spec, plant_spec2)
x3 <- bind_rows(damspec2,damspec)
x4 <- bind_rows(valvespec2,valvespec)
x5 <- bind_rows(gatespec, gatespec2)

x <- x1 %>% full_join(x2) %>% full_join(x3) %>% full_join(x4) %>% full_join(x5)

```


```{r}
data1 <- data.frame(data1=runif(10),data2=runif(10),data4=runif(10))
data2 <- data.frame(data1=runif(10),data3=runif(10),data5=runif(10))

datatotal <- bind_rows(data1,data2)
```

